---
title: "poll_prediction"
author: "Karim Toufiq"
date: "2024-03-08"
output: html_document
---


```{r setup, include=TRUE}
# BOOTSTRAP EXAMPLE
# Load necessary library
library(boot)

# Assume 'data' is your data vector containing numeric values
# For example, let's create some sample data
set.seed(123) # For reproducibility
data <- rnorm(100, mean = 50, sd = 10) # 100 random normal numbers with mean 50 and sd 10

# Define a function to obtain the mean of a sample
mean_function <- function(data, indices) {
  # The 'indices' parameter allows the boot function to select bootstrap samples
  return(mean(data[indices]))
}

# Perform the bootstrap
set.seed(123) # For reproducibility
bootstrap_results <- boot(data = data, statistic = mean_function, R = 1000)

# Calculate the 95% confidence interval
bootstrap_ci <- boot.ci(bootstrap_results, type = "perc")

# Print the bootstrap estimate and confidence interval
print(bootstrap_results$t0)
print(bootstrap_ci)
```

Notes for Bootstrapping Polling Data
Following: 
https://towardsdatascience.com/bootstrapping-method-to-estimate-voters-preferences-in-us-presidential-election-2020-d480ef13aef

```{python, include=FALSE}
# IMPORT PACKAGES FOR PYTHON
import pandas as pd
#LOAD DATA
polling = pd.read_csv("https://raw.githubusercontent.com/PriyaBrataSen/US-Election-Poll/main/pres_polls.csv")
```

```{r, include=TRUE}
library(reticulate)
library(dplyr)
library(boot)

# Store Python Object as a R Object
df <- as.data.frame(py$polling) # Saves Python Object as R

# Extract unique pollster names from the 'Pollster' column of the dataframe 'df'
poll = unique(pull(df, Pollster))

# Display the structure of the 'poll' object to understand its contents and type
str(poll)

# Define a function 'bootmean' to calculate the mean of a subset of data
bootmean = function(x, i) { mean(x[i]) }

# Define a function 'prefer_country' to perform bootstrap analysis on a given dataset
prefer_country = function(data) {
  # Perform bootstrap resampling 10,000 times using the 'bootmean' function
  boot.object = boot(data, bootmean, R = 10000)

  # Calculate the 95% confidence interval using bias-corrected accelerated (BCa) method
  boot.ci(boot.object, conf = 0.95, type = 'bca')
}

# Calculate the confidence interval for the 'Dem' variable and round to 4 decimal places
Dem = round(prefer_country(df$Dem)$bca[, c(4, 5)], 4)

# Calculate the confidence interval for the 'GOP' variable and round to 4 decimal places
GOP = round(prefer_country(df$GOP)$bca[, c(4, 5)], 4)

# Create a vector with the Democratic party's confidence interval
c('Democratic party:', Dem)

# Create a vector with the Republican party's confidence interval
c('Republican party:', GOP)

# Initialize empty vectors for storing lower and upper confidence limits and region names
lower = c()
upper = c()
region = c()

# Extract unique region names from the 'Region' column of the dataframe 'df'
a = unique(pull(df, Region))

# Define a function 'prefer_region' to perform region-based bootstrap analysis
prefer_region = function(data) {
  for (i in a) {
    # Subset data for each region
    data_Dem = data[df$Region == i]

    # Perform bootstrap analysis on the subset data
    boot.Dem = boot(data_Dem, bootmean, R = 10000)

    # Calculate the 95% confidence interval for the subset data
    p = boot.ci(boot.Dem, conf = 0.95)

    # Store the lower and upper bounds and region names in their respective vectors
    lower = c(lower, p$bca[, c(4)])
    upper = c(upper, p$bca[, c(5)])
    region = c(region, i)
  }

  # Combine the region, lower, and upper bounds into a dataframe
  preference = data.frame(region, lower, upper)
  preference
}

# Calculate the confidence intervals for each region for the Democratic party
DEM = prefer_region(df$Dem) %>% rename(Dem_lower = lower, Dem_upper = upper)

# Calculate the confidence intervals for each region for the Republican party
GOP = prefer_region(df$GOP) %>% rename(GOP_lower = lower, GOP_upper = upper)

# Merge the Democratic and Republican party dataframes by region
inner_join(DEM, GOP, by = 'region')
```

```{r, include=TRUE}
# Define the regression function for bootstrapping
regression_function <- function(data, indices) {
  # Sample the data
  d <- data[indices, ]  # Bootstrap sample
  
  # Fit the model
  fit <- lm(Dem ~ Day + as.factor(Region), data = d)
  
  # Return the coefficients of the model
  # If the model fails for a sample, return a vector of NAs
  if (length(coef(fit)) == length(coefficients(lm(Dem ~ Day + as.factor(Region), data = data)))) {
    return(coef(fit))
  } else {
    return(rep(NA, length(coefficients(lm(Dem ~ Day + as.factor(Region), data = data)))))
  }
}

# Assuming 'df' is your dataframe and it's already been loaded into your R environment.
# Perform the bootstrap
set.seed(123) # For reproducibility
boot_results_Dem <- boot(data = df, statistic = regression_function, R = 1000)

# Summarize the bootstrap results
print(boot_results_Dem)

# Obtain bootstrap confidence intervals for the first coefficient (intercept) as an example
boot_ci_Dem <- boot.ci(boot_results_Dem, type = "perc", index = 1)
print(boot_ci_Dem)
```

# Make a Forecast on that Bootstrap

```{r, include=TRUE}
# Fit the linear regression model to the existing data
fit <- lm(Dem ~ Day + as.factor(Region), data = df)

# Make sure 'Region' in 'df' is a factor
df$Region <- as.factor(df$Region)

# Create new data frame for future days
max_day <- max(df$Day)
future_days <- data.frame(Day = seq(from = max_day + 1, by = 1, length.out = 100))

# Assign the first level of 'Region' from 'df' to 'future_days'
# Make sure that the 'Region' variable has levels
if(length(levels(df$Region)) > 0) {
  future_days$Region <- rep(levels(df$Region)[1], times = 100)
} else {
  stop("No levels found in df$Region. Ensure it is a factor with levels.")
}

# Predict future 'Dem' values
future_predictions <- predict(fit, newdata = future_days)

# Add the predictions as a new column in 'future_days'
future_days$Dem <- future_predictions

# Prepare 'df' and 'future_days' for binding by selecting matching columns
df <- df[, c("Day", "Dem")]
future_days <- future_days[, c("Day", "Dem")]

# Combine historical and forecasted data
combined_data <- rbind(df, future_days)

# Calculate base standard error from the model
se_fit <- sqrt(sum(residuals(fit)^2) / fit$df.residual)

# Add a time distance measure to future_days (distance from the last observed day)
future_days$time_distance <- seq_along(future_days$Day)

# Modify the scaling factor for the standard error to reduce the spread
# For example, using a smaller exponent or a constant less than 1
scaling_factor <- 0.025  # Adjust this value as needed
future_days$se <- se_fit * (sqrt(future_days$time_distance) * scaling_factor)

# Calculate upper and lower bounds for confidence intervals
future_days$upper <- future_days$Dem + 1.96 * future_days$se
future_days$lower <- future_days$Dem - 1.96 * future_days$se
library(ggplot2)
# Plot
ggplot() +
  geom_line(data = df, aes(x = Day, y = Dem), color = "#3f8fc4", alpha = 0.5) +
  geom_point(data = df, aes(x = Day, y = Dem), color = "#3f8fc4", alpha = 0.1) +
  geom_line(data = future_days, aes(x = Day, y = Dem), color = "#08306b", linetype = "dashed") +
  geom_ribbon(data = future_days, aes(x = Day, ymin = lower, ymax = upper), 
              fill = "#3f8fc4", alpha = 0.2) +
  theme_minimal() +
  labs(title = "Democrat Support Over Time with Forecast",
       x = "Day",
       y = "Democrat Support")

```

From: https://slate.com/news-and-politics/2018/03/how-much-do-democrats-need-to-win-by.html

"It took a 5.4-percentage-point lead in the national congressional vote for Democrats to gain the 31 seats they won in the last Democratic wave in 2006. Republicans held over 55 percent of the seats in the House after 2016’s elections despite winning only 49.9 percent of the popular vote."

```{r setup, include=TRUE, warning=FALSE}
# Filter out the initial volatile values, for example, the first 10 days
df_filtered <- df %>% filter(Day > 10)

# Define a y-axis range that excludes extreme values, for example from 30 to 70
y_axis_range <- c(30, 70)

# Plot with the filtered data and adjusted y-axis range
library(ggplot2)

ggplot() +
  geom_line(data = combined_data, aes(x = Day, y = Dem), color = "#3f8fc4", alpha = 0.5) +
  geom_point(data = df_filtered, aes(x = Day, y = Dem), color = "#3f8fc4", alpha = 0.1) +
  geom_line(data = future_days, aes(x = Day, y = Dem), color = "#08306b", linetype = "dashed") +
  geom_ribbon(data = future_days, aes(x = Day, ymin = lower, ymax = upper), 
              fill = "#08306b", alpha = 0.2) +
  geom_hline(yintercept = 55.4, color = "black", linetype = "solid", size = 0.5) +
  theme_minimal() +
  labs(title = "Democrat Support Over Time with Forecast",
       x = "Day",
       y = "Democrat Support") +
  scale_y_continuous(limits = y_axis_range)  # Set y-axis limits
```

# PREDITION BASED ON 356 DAYS

```{r setup, include=TRUE, warning=FALSE}
# RESTART 

# Store Python Object as a R Object
df <- as.data.frame(py$polling) # Saves Python Object as R

# Extract unique pollster names from the 'Pollster' column of the dataframe 'df'
poll = unique(pull(df, Pollster))

# Display the structure of the 'poll' object to understand its contents and type
str(poll)

# Define a function 'bootmean' to calculate the mean of a subset of data
bootmean = function(x, i) { mean(x[i]) }

# Define a function 'prefer_country' to perform bootstrap analysis on a given dataset
prefer_country = function(data) {
  # Perform bootstrap resampling 10,000 times using the 'bootmean' function
  boot.object = boot(data, bootmean, R = 10000)

  # Calculate the 95% confidence interval using bias-corrected accelerated (BCa) method
  boot.ci(boot.object, conf = 0.95, type = 'bca')
}

# Calculate the confidence interval for the 'Dem' variable and round to 4 decimal places
Dem = round(prefer_country(df$Dem)$bca[, c(4, 5)], 4)

# Calculate the confidence interval for the 'GOP' variable and round to 4 decimal places
GOP = round(prefer_country(df$GOP)$bca[, c(4, 5)], 4)

# Create a vector with the Democratic party's confidence interval
c('Democratic party:', Dem)

# Create a vector with the Republican party's confidence interval
c('Republican party:', GOP)

# Initialize empty vectors for storing lower and upper confidence limits and region names
lower = c()
upper = c()
region = c()

# Extract unique region names from the 'Region' column of the dataframe 'df'
a = unique(pull(df, Region))

# Define a function 'prefer_region' to perform region-based bootstrap analysis
prefer_region = function(data) {
  for (i in a) {
    # Subset data for each region
    data_Dem = data[df$Region == i]

    # Perform bootstrap analysis on the subset data
    boot.Dem = boot(data_Dem, bootmean, R = 10000)

    # Calculate the 95% confidence interval for the subset data
    p = boot.ci(boot.Dem, conf = 0.95)

    # Store the lower and upper bounds and region names in their respective vectors
    lower = c(lower, p$bca[, c(4)])
    upper = c(upper, p$bca[, c(5)])
    region = c(region, i)
  }

  # Combine the region, lower, and upper bounds into a dataframe
  preference = data.frame(region, lower, upper)
  preference
}

# Calculate the confidence intervals for each region for the Democratic party
DEM = prefer_region(df$Dem) %>% rename(Dem_lower = lower, Dem_upper = upper)

# Calculate the confidence intervals for each region for the Republican party
GOP = prefer_region(df$GOP) %>% rename(GOP_lower = lower, GOP_upper = upper)

# Merge the Democratic and Republican party dataframes by region
inner_join(DEM, GOP, by = 'region')

# Define the regression function for bootstrapping
regression_function <- function(data, indices) {
  # Sample the data
  d <- data[indices, ]  # Bootstrap sample
  
  # Fit the model
  fit <- lm(Dem ~ Day + as.factor(Region), data = d)
  
  # Return the coefficients of the model
  # If the model fails for a sample, return a vector of NAs
  if (length(coef(fit)) == length(coefficients(lm(Dem ~ Day + as.factor(Region), data = data)))) {
    return(coef(fit))
  } else {
    return(rep(NA, length(coefficients(lm(Dem ~ Day + as.factor(Region), data = data)))))
  }
}

# Assuming 'df' is your dataframe and it's already been loaded into your R environment.
# Perform the bootstrap
set.seed(123) # For reproducibility
boot_results_Dem <- boot(data = df, statistic = regression_function, R = 1000)

# Summarize the bootstrap results
print(boot_results_Dem)

# Obtain bootstrap confidence intervals for the first coefficient (intercept) as an example
boot_ci_Dem <- boot.ci(boot_results_Dem, type = "perc", index = 1)
print(boot_ci_Dem)

fit <- lm(Dem ~ Day + as.factor(Region), data = df)

# Make sure 'Region' in 'df' is a factor
df$Region <- as.factor(df$Region)

# Create new data frame for future days
max_day <- max(df$Day)
future_days <- data.frame(Day = seq(from = max_day + 1, by = 1, length.out = 356))

# Assign the first level of 'Region' from 'df' to 'future_days'
# Make sure that the 'Region' variable has levels
if(length(levels(df$Region)) > 0) {
  future_days$Region <- rep(levels(df$Region)[1], times = 356)
} else {
  stop("No levels found in df$Region. Ensure it is a factor with levels.")
}

# Predict future 'Dem' values
future_predictions <- predict(fit, newdata = future_days)

# Add the predictions as a new column in 'future_days'
future_days$Dem <- future_predictions

# Prepare 'df' and 'future_days' for binding by selecting matching columns
df <- df[, c("Day", "Dem")]
future_days <- future_days[, c("Day", "Dem")]

# Combine historical and forecasted data
combined_data <- rbind(df, future_days)

# Calculate base standard error from the model
se_fit <- sqrt(sum(residuals(fit)^2) / fit$df.residual)

# Add a time distance measure to future_days (distance from the last observed day)
future_days$time_distance <- seq_along(future_days$Day)

# Modify the scaling factor for the standard error to reduce the spread
# For example, using a smaller exponent or a constant less than 1
scaling_factor <- 0.025  # Adjust this value as needed
future_days$se <- se_fit * (sqrt(future_days$time_distance) * scaling_factor)

# Calculate upper and lower bounds for confidence intervals
future_days$upper <- future_days$Dem + 1.96 * future_days$se
future_days$lower <- future_days$Dem - 1.96 * future_days$se

# Filter out the initial volatile values, for example, the first 10 days
df_filtered <- df %>% filter(Day > 10)

# Define a y-axis range that excludes extreme values, for example from 30 to 70
y_axis_range <- c(30, 70)

# Plot with the filtered data and adjusted y-axis range
ggplot() +
  geom_line(data = combined_data, aes(x = Day, y = Dem), color = "#3f8fc4", size = 0.8) +
  geom_point(data = df_filtered, aes(x = Day, y = Dem), color = "#3f8fc4", alpha = 0.6, size = 1.5) +
  geom_line(data = future_days, aes(x = Day, y = Dem), color = "#08306b", linetype = "dashed", size = 0.8) +
  geom_ribbon(data = future_days, aes(x = Day, ymin = lower, ymax = upper), 
              fill = "#08306b", alpha = 0.2) +
  geom_hline(yintercept = 55.4, color = "black", linetype = "solid", size = 0.5) +
  theme_minimal() +
  labs(
    title = "Forecast of Democrat Support Over Time",
    subtitle = "With 95% Confidence Intervals",
    x = "Day",
    y = "Democrat Support (%)",
  ) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "gray80"),
    panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "gray90")
  ) +
  scale_y_continuous(limits = y_axis_range)  # Set y-axis limits
```

